### 2018-01-07 Writing/Thesis/content/chapter-appendix.tex
Add introduction, related + fundamentals parts

% !TEX root = ../my-thesis.tex
%
\chapter{Example Appendix}
\label{sec:appendix}

\section{Appendix Section 1}
\label{sec:appendix:sec1}

\begin{table}[h]
	\begin{tabularx}{\textwidth}{X | X | X}
		%\hline
		Alpha		& Beta			& Gamma			\\ \hline
		0			& 1				& 2				\\ \hline
		3			& 4				& 5				\\ %\hline
	\end{tabularx}
	\label{tab:table1}
	\caption{This is a caption text.}
\end{table}