### 2018-01-07 Writing/Thesis/content/chapter-implementation.tex

% !TEX root = ../my-thesis.tex
%
\chapter{Implementation}
\label{sec:implementation}