### 2018-01-14 Writing/Thesis/content/chapter-implementation.tex
Write some more for thesis

% !TEX root = ../my-thesis.tex
%
\chapter{Implementation}
\label{sec:implementation}

\subsection{Frameworks used}
% Weka for classifiers

% Data sets from OpenML

\subsection{jPL framework}
The jPL framework is a java framework for the evaluation of preference learning algorithms. It implements several tasks from the context of preference learning, including label ranking. Furthermore, the framework introduces the Generic Preference Representation Format (GPRF) to store preference information.

\subsection{WEKA}