### 2018-02-06 Writing/Thesis/content/abstract.tex
more writing

% !TEX root = ../my-thesis.tex
%
\pdfbookmark[0]{Abstract}{Abstract}
\chapter*{Abstract}
\label{sec:abstract}
\vspace*{-10mm}

Should abstract be included?

% Motivation

% Problem Statement

% Approach

% Results

% Conclusion

\vspace*{20mm}

{\usekomafont{chapter}Abstract (German)}\label{sec:abstract-german} \\

Und falls ja, dann zweisprachig?

\textbf{Keywords: } Automated Machine Learning, Meta Learning, Regression, Label Ranking, Algorithm Ranking, Algorithm Selection