### 2018-02-06 Writing/Thesis/content/chapter-implementation.tex

% !TEX root = ../my-thesis.tex
%
\chapter{Implementation}
\label{sec:implementation}

\subsection{Frameworks used}
% Weka for classifiers

% Data sets from OpenML

\subsection{jPL framework}
The jPL framework is a java framework for the evaluation of preference learning algorithms. It implements several tasks from the context of preference learning, including label ranking. Furthermore, the framework introduces the Generic Preference Representation Format (GPRF) to store preference information.

\subsection{WEKA}

% openml Webapp for metaFeature implementation except default accuracy

\section{Generating Performance Values}
Performance values generated by

\section{Stuff}

Rankers trained with data set of the following form, so each ranking algorithm can derive its on performance