### 2018-02-12 Writing/Thesis/content/chapter-appendix.tex

% !TEX root = ../my-thesis.tex
%
\chapter{Appendix}
\label{sec:appendix}

\begin{figure}
\includegraphics[width=\textwidth]{gfx/placeholder_allMetaFeatures}
\caption{PLACEHOLDER: The meta features computed by each meta feature group. The full list of meta features encompasses all features enumerated in the right column.}
\label{fig:metaFeaturesDetails}
\end{figure}