### 2018-03-14 Writing/Thesis/Tex/content/acknowledgement.tex
Create Presentation

\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}

I would like to thank Prof. Dr. Hüllermeier for trusting me with this topic, and for his useful feedback on the matter. A very special gratitude goes out to Marcel Wever and Felix Mohr, without whose helpful insights and encouragement this thesis could not have been completed. Thank you for your explanations when I was struggling, and for pointing me in the right direction when I ran the risk of losing focus. My sincere thanks also goes to the proofreaders, who meticulously corrected my spelling and pointed out weak explanations. Last but not least, I am grateful for my family and friends for always supporting me and believing in me, and reminding me that I, too, need some sleep once in a while.